---
title: "Untitled"
output: html_document
---


#Loading and preprocessing the data
```{r echo=TRUE}
setwd("C:/GDrive/Coursera/ReproducibleResearch/Peer1")

rawdata <-read.csv("activity.csv")

```


#What is mean total number of steps taken per day?
##1.Calculate the total number of steps taken per day

Removing NA Values from this point forward
```{r, echo=TRUE}
StepsADay <-aggregate(rawdata$steps,by=list(rawdata$date),FUN=sum,na.rm=TRUE)
#Rename Columns
names(StepsADay)<-c("date","totalsteps")
StepsADay
```

##2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day


```{r, echo=TRUE}
hist(StepsADay$totalsteps,
     xlab = "Total Number of Steps",
     main = " Histogram of Number of Steps per Day")
```

##3. Calculate and report the mean and median of the total number of steps taken per day

mean
```{r, echo=TRUE}
mean(StepsADay$totalsteps)

```

median

```{r echo=TRUE}
median(StepsADay$totalsteps)
```

#What is the average daily activity pattern?

##1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r echo=TRUE}
AvgStepsbyInterval<-aggregate(rawdata$steps, by=list(rawdata$interval), FUN=mean,na.rm=TRUE)
#Rename Columns
names(AvgStepsbyInterval)<-c("interval","totalsteps")

plot(AvgStepsbyInterval$interval,
     AvgStepsbyInterval$totalsteps,
     type="l",
     xlab="Interval, minutes",
     ylab="Avg Steps")
```

##2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}
#find the interval
AvgStepsbyInterval[which(AvgStepsbyInterval$totalsteps == max(AvgStepsbyInterval$totalsteps)),]
```


#Inputing missing values

##1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r echo=TRUE}
sum(is.na(rawdata$steps))
```



##2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r echo=TRUE}
#replace na with mean of all intervals

mean(rawdata$steps,na.rm=TRUE)

```

##3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r echo=TRUE}
#copy rawdata
mydata<-rawdata

#replace na with mean of all intervals
mydata[is.na(mydata)]<-mean(rawdata$steps,na.rm=TRUE)
```


##4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r echo=TRUE}

revStepsADay <-aggregate(mydata$steps,by=list(mydata$date),FUN=sum,na.rm=TRUE)
#Rename Columns
names(revStepsADay)<-c("date","totalsteps")
StepsADay

hist(revStepsADay$totalsteps,
     xlab = "Total Number of Steps",
     main = " Histogram of Number of Steps per Day")

mean(revStepsADay$totalsteps)

median(revStepsADay$totalsteps)

```
###Do these values differ from the estimates from the first part of the assignment? 
yes


###What is the impact of imputing missing data on the estimates of the total daily number of steps?
A slight increase on the mean, not much affect on median

#Are there differences in activity patterns between weekdays and weekends?

##1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r echo=TRUE}

library("chron")

mydata$weekend = is.weekend(mydata$date)
temp<-mydata

temp$DayType <-ifelse(temp$weekend == FALSE, "WEEKDAY","WEEKEND")

```

##2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r echo=TRUE}
library("lattice")
temp2<-aggregate(temp$steps, by=list(temp$interval,temp$DayType), FUN=mean,na.rm=TRUE)

names(temp2)<-c("interval","daytype","mean")

xyplot(mean ~ interval | daytype, temp2,
       type="l",
       xlab="Interval", 
       ylab="Number of steps", 
       layout=c(1,2))
```
